---
title: "categorical variable prediction"
output: html_notebook
---

[section 5.13.4](https://jhudatascience.org/tidyversecourse/model.html#example-of-categorical-variable-prediction) from tidymodels ecosystems - chapter 5 of _ts4ds_

```{r}
library(tidymodels)
```



We have already split our data into testing and training sets (see 01_LEARNING.Rmd), so we don’t necessarily need to do that again.

However, we can stratify our split by a particular feature of the data using the strata argument of the initial_split() function.


```{r}
set.seed(1234)
initial_split(iris, strata = Species, prop = 2/3)
```

```{r}
training_iris <-training(split_iris)
training_iris
```

```{r}
count(training_iris, Species)
```

```{r}
testing_iris <-testing(split_iris)
testing_iris
```


```{r}
count(testing_iris, Species)
```

this time we'll perform cross-validation

Cross validation is also helpful for optimizing what we call hyperparameters.

Hyperparameters are aspects about the model that we need to specify. 


Often packages will choose a default value, however it is better to use the training data to see what value





