---
title: "Build a model"
output: 
  html_document:
    toc: true
---

## separate what you want to do from the execution

"The extra step of defining the model using a function like `linear_reg()` might seem superfluous since a call to `lm()` is _much more succinct_. However, the problem with standard modeling functions is that they don’t **separate what you want to do from the execution**. For example, the process of executing a formula has to happen repeatedly across model calls even when the formula does not change; we can’t **recycle those computations**.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
options(tibble.print_min = 5)
```

Get started with building a model in this R Markdown document that accompanies [Build a model](https://www.tidymodels.org/start/models/) tidymodels start article.

If you ever get lost, you can visit the links provided next to section headers to see the accompanying section in the online article.

Take advantage of the RStudio IDE and use "Run All Chunks Above" or "Run Current Chunk" buttons to easily execute code chunks. If you have been running other tidymodels articles in this project, restart R before working on this article so you don't run out of memory on RStudio Cloud.


## [Introduction](https://www.tidymodels.org/start/models/#intro)

Load necessary packages:

```{r}
library(tidyverse)
library(tidymodels)  # for the parsnip package, along with the rest of tidymodels

# Helper packages
library(readr)       # for importing data
library(broom.mixed) # for converting bayesian models to tidy tibbles
```


## [The Sea Urchins Data](https://www.tidymodels.org/start/models/#data)


```{r}
urchins <-
  # Data were assembled for a tutorial 
  # at https://www.flutterbys.com.au/stats/tut/tut7.5a.html
  read_csv("https://tidymodels.org/start/models/urchins.csv") %>% 
  # Change the names to be a little more verbose
  setNames(c("food_regime", "initial_volume", "width")) %>% 
  # Factors are very helpful for modeling, so we convert one column
  mutate(food_regime = factor(food_regime, levels = c("Initial", "Low", "High")))
```

Look at the data:

```{r}
urchins
```

Plot the data: 

```{r}
ggplot(urchins,
       aes(x = initial_volume, 
           y = width, 
           group = food_regime, 
           col = food_regime)) + 
  geom_point() + 
  geom_smooth(method = lm, se = FALSE) +
  scale_color_viridis_d(option = "plasma", end = .7)
```

## [Build and fit a model](https://www.tidymodels.org/start/models/#build-model)

```{r}
linear_reg() %>% 
  set_engine("lm")

```

Try typing `?linear_reg()` in the console to see all available engines and other details about this model type.

Create model specification:

```{r}
lm_mod <- 
  linear_reg() %>% 
  set_engine("lm")
```

> For this model, other packages may add additional engines. Use `show_engines()` to see the current set of engines. e.g. `show_engines("linear_reg")`


Fit model:

```{r}
lm_fit <- 
  lm_mod %>% 
  fit(width ~ initial_volume * food_regime, data = urchins)
lm_fit
```

Present model results in a tidyverse friendly way with `tidy()` from `broom` package.

```{r}
tidy(lm_fit)
```

This kind of output can be used to generate a dot-and-whisker plot of our regression results using the dotwhisker package:


```{r}
tidy(lm_fit) %>% 
  dotwhisker::dwplot(dot_args = list(size = 2, color = "black"),
         whisker_args = list(color = "black"),
         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
```



## [Use a model to predict](https://www.tidymodels.org/start/models/#predict-model)

New example data to predict:

```{r}
new_points <- expand.grid(initial_volume = 20, 
                          food_regime = c("Initial", "Low", "High"))
new_points
```

```{r}
urchins %>% 
  filter(initial_volume == 20)
```


Generate the mean body width values: 

```{r}
mean_pred <- predict(lm_fit, new_data = new_points)
mean_pred
```

Get confidence intervals and plot:

```{r}
conf_int_pred <- predict(lm_fit, 
                         new_data = new_points, 
                         type = "conf_int")
conf_int_pred

# Now combine: 
plot_data <- 
  new_points %>% 
  bind_cols(mean_pred) %>% 
  bind_cols(conf_int_pred)

# and plot:
ggplot(plot_data, aes(x = food_regime)) + 
  geom_point(aes(y = .pred)) + 
  geom_errorbar(aes(ymin = .pred_lower, 
                    ymax = .pred_upper),
                width = .2) + 
  labs(y = "urchin size")
```

## [Model with a different engine](https://www.tidymodels.org/start/models/#new-engine)

Switch to Bayesian approach by simply changing your engine to **stan**: 

```{r}
# set the prior distribution
prior_dist <- rstanarm::student_t(df = 1)

set.seed(123)

# make the parsnip model
bayes_mod <-   
  linear_reg() %>% 
  set_engine("stan", 
             prior_intercept = prior_dist, 
             prior = prior_dist) 

# train the model
bayes_fit <- 
  bayes_mod %>% 
  fit(width ~ initial_volume * food_regime, data = urchins)

print(bayes_fit, digits = 5)
```

To update the parameter table, the `tidy()` method is once again used: 

```{r}
tidy(bayes_fit, intervals = TRUE)
```

Get your predictions without changing the syntax you used earlier:

```{r}
bayes_plot_data <- 
  new_points %>% 
  bind_cols(predict(bayes_fit, new_data = new_points)) %>% 
  bind_cols(predict(bayes_fit, new_data = new_points, type = "conf_int"))

ggplot(bayes_plot_data, aes(x = food_regime)) + 
  geom_point(aes(y = .pred)) + 
  geom_errorbar(aes(ymin = .pred_lower, ymax = .pred_upper), width = .2) + 
  labs(y = "urchin size") + 
  ggtitle("Bayesian model with t(1) prior distribution")
```


Think about how we are using the pipe (`%>%`):

+ Use the pipe to pass around the _data_ in the **tidyverse**   
+ Use the pipe to pass around the _model object_ with **tidymodels**


### Applying to previous knowledge

From an example found at [_OpenIntro_](https://openintro-ims.netlify.app/)

```{r}
library(openintro)
```



```{r}
loans <- loans_full_schema %>% 
  select(interest_rate, debt_to_income, 
         term, issue_month, verified_income,
         total_credit_utilized, total_credit_limit,
         public_record_bankrupt, 
         credit_checks = inquiries_last_12m) %>% 
  mutate(credit_util = total_credit_utilized / total_credit_limit) %>% 
  mutate(bankruptcy = case_when(
    public_record_bankrupt == 2 ~ 1,
    public_record_bankrupt == 3 ~ 1,
    TRUE ~ as.numeric(public_record_bankrupt)
  )) %>% 
  relocate(credit_util, .after = debt_to_income) %>% 
  relocate(bankruptcy, .after = credit_util) %>% 
  select(-total_credit_utilized, -total_credit_limit, -public_record_bankrupt) 

# loans$bankruptcy
# bankrupt_chrfct <- recode(loans$bankruptcy, "0" = "no", "1" = "yes")
# bankrupt_chrfct
# 
# loans$bankrupt_chrfct <- bankrupt_chrfct

loans
```

1. specify the model: `linear_reg()`
2. identify the engine:  `set_engine("lm")`
3. train or estimate the model (fit): `fit(interest_rate ~ bankruptcy, data = loans)`  

    - `show_engines("linear_reg")`

```{r}
show_engines("linear_reg")
# fit_credit = lm(interest_rate ~ bankruptcy, data = loans)

lm_fit_credit <- linear_reg() %>% 
  set_engine("lm") %>% 
  fit(interest_rate ~ bankruptcy, data = loans)
lm_fit_credit
```

```{r}
tidy(lm_fit_credit)
glance(lm_fit_credit)
```


```{r}
lm_fit_multiple <- linear_reg() %>%
  set_engine("lm") %>%
  fit(
    interest_rate ~ verified_income + 
      debt_to_income + 
      credit_util + bankruptcy + term + 
      credit_checks + issue_month,
    data = loans
  )

glance(lm_fit_multiple)
tidy(lm_fit_multiple)
```


```{r}
loans %>% 
  ggplot(aes(credit_checks, interest_rate,
         col = verified_income)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE) +
  scale_color_viridis_d(option = "plasma", end = .7)
```

```{r}
library(dotwhisker)
tidy(lm_fit_multiple) %>% 
  dwplot(dot_args = list(size = 2, color = "black"),
         whisker_args = list(color = "black"),
         vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2))
```


```{r}
urchins

loans

expand.grid()
```

